---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# `R/rw`: Robins-Wang variance estimation for multiple imputation

<!-- badges: start -->
[![R-CMD-check](https://github.com/LucyMcGowan/rw/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/LucyMcGowan/rw/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/LucyMcGowan/rw/graph/badge.svg)](https://app.codecov.io/gh/LucyMcGowan/rw)
<!-- badges: end -->

The purpose of the `rw` package is to compute [Robins-Wang variance estimates](https://doi.org/10.1093/biomet/87.1.113) for multiply imputed data analyses. The core workflow currently supports `mice` imputation models `method = "norm"` and `method = "logreg"`. For predictive mean matching (PMM), use `method = "pmmrw"`, a PMM imputer supplied by this package that records the observed donor row used for each imputed recipient. Those donor IDs allow `pool_rw()` to apply a PMM donor-source covariance adjustment to the analysis-score component.

## Installation

This package requires the development version of mice with the `tasks` argument:

```r
remotes::install_github("amices/mice@dev")
```

Then you can install the development version of `rw` like so:

``` r
remotes::install_github("LucyMcGowan/rw")
```


## Parametric imputation example

The first step is to impute your data using `mice`. When using the `mice` function, you must set the parameter `tasks = "train"` in order to pass the necessary parts to our subsequent functions. In this example, we use the `nhanes` data from the mice package and impute the incomplete variables `bmi`, `hyp`, and `chl` using normal working models. The variable `age` has no missing values.

```{r}
#| message: false
library(rw)
library(mice)
set.seed(1)
nhanes_scaled <- as.data.frame(scale(nhanes))

meth_norm <- make.method(nhanes_scaled)
meth_norm[] <- ""
meth_norm[c("bmi", "hyp", "chl")] <- "norm"

imp_norm <- mice(nhanes_scaled, 
                 method = meth_norm,
                 m = 5, 
                 tasks = "train",
                 print = FALSE)
```

Now, suppose we want to fit a model predicting `bmi` from `age` and `hyp`. The `with_rw` function allows this, where the first argument is the imputation object from the `mice` function above and the second is the model expression. Note that currently outcome models must be Gaussian or Binomial.

```{r}
fit_rw_norm <- with_rw(imp_norm, lm(bmi ~ age + hyp))
```

Finally, we can pool the results from the fit object and calculate the Robins-Wang variance using the `pool_rw` function

```{r}
pool_rw(fit_rw_norm)
```

Let's compare this result to using Rubin's rules.

```{r}
fit_rr_norm <- with(imp_norm, lm(bmi ~ age + hyp))
pool(fit_rr_norm) |>
  summary()
```


## PMM imputation with donor-source correlation

For PMM, several imputed rows may use the same observed donor value. Standard `mice` PMM does not save the realized donor row IDs, so `rw` provides `method = "pmmrw"` to record those IDs during imputation. In the donor ID matrix, `donor_id[i, p]` is `NA` if row `i` was not imputed by PMM in imputation `p`; otherwise it is the row number of the observed donor used for row `i`.

When donor IDs are available, `pool_rw()` applies the PMM donor-source
correlation adjustment to the analysis-score covariance term. Let
\(U_i^{(p)}\) denote the complete-data analysis score contribution for row
\(i\) in imputation \(p\), and let \(J_i^{(p)}\) be the observed donor row used
for PMM recipient \(i\). For rows that were not PMM recipients,
\(J_i^{(p)} = \texttt{NA}\).

For each imputation, `pool_rw()` forms a donor-source score matrix
\(\widetilde U^{(p)}\) by adding each PMM recipient score contribution to the
row of its observed donor:

$$\widetilde U_j^{(p)} = U_j^{(p)} + \sum_{i: J_i^{(p)} = j} U_i^{(p)}, \qquad \widetilde U_i^{(p)} = 0 \quad\text{if } J_i^{(p)} \text{ is observed}.$$


The analysis-score covariance term is then computed from the donor-source
scores,

$$\Omega_{\mathrm{PMM}} = \frac{1}{n}\left(\frac{1}{m}\sum_{p=1}^m \widetilde U^{(p)}\right)^\top\left(\frac{1}{m}\sum_{p=1}^m \widetilde U^{(p)}\right).$$

rather than from the ordinary rowwise scores. This adjustment captures the
correlation induced when several PMM recipients copy values from the same
observed donor.

The Gaussian working model in `pmmrw` is used to construct the PMM matching
index; the copied PMM value is not treated as a Gaussian draw in the
Robins-Wang nuisance-score terms. Regular `norm` and `logreg` imputation
blocks, if present, still contribute their usual Robins-Wang score components.

Here is a PMM example. We impute `bmi` by PMM using complete predictor `age`, then analyze `bmi ~ age`.

```{r}
meth_pmm <- make.method(nhanes_scaled)
meth_pmm[] <- ""
meth_pmm["bmi"] <- "pmmrw"

meth_std_pmm <- meth_pmm
meth_std_pmm["bmi"] <- "pmm"

pred_pmm <- make.predictorMatrix(nhanes_scaled)
pred_pmm[,] <- 0
pred_pmm["bmi", "age"] <- 1
```

Using the same random seed, standard `mice` PMM and `pmmrw` produce the same imputed values. The difference is that `pmmrw` also stores donor row IDs.

```{r}
set.seed(1)
imp_std_pmm <- mice(nhanes_scaled,
                    method = meth_std_pmm,
                    predictorMatrix = pred_pmm,
                    m = 5,
                    tasks = "train",
                    print = FALSE)

set.seed(1)
imp_pmm <- mice(nhanes_scaled,
                method = meth_pmm,
                predictorMatrix = pred_pmm,
                m = 5,
                tasks = "train",
                print = FALSE)

identical(imp_std_pmm$imp$bmi, imp_pmm$imp$bmi)

donor_id <- extract_donor_id(imp_pmm, "bmi")
head(donor_id)
```

```{r}
fit_rw_pmm <- with_rw(imp_pmm, lm(bmi ~ age))

# donor_id is extracted automatically for method = "pmmrw";
# pool_rw() then applies the donor-source covariance adjustment.
pool_rw(fit_rw_pmm)
```

Compare this with Rubin's rules for the same PMM imputations.

```{r}
fit_rr_pmm <- with(imp_pmm, lm(bmi ~ age))
pool(fit_rr_pmm) |>
  summary()
```

## Giganti & Shepherd Example


Below is an example to replicate the [Giganti & Shepherd (2020)](https://doi.org/10.1093/aje/kwaa153) results. This example uses the parametric `norm` and `logreg` imputation models and demonstrates the original Robins-Wang workflow; it is not a PMM correlation-adjusted example.

```{r example}
set.seed(1)
meth <- make.method(giganti_data)
meth[] <- ""
meth[c("A", "D")] <- c("norm", "logreg")

pred <- make.predictorMatrix(giganti_data)
pred[,] <- 0
pred["A", c("X1", "X2", "A.star", "D.star")] <- 1
pred["D", c("X1", "X2", "A.star", "D.star", "A")] <- 1

imp <- mice(giganti_data, m = 10, method = meth, predictorMatrix = pred,
            print = FALSE, tasks = "train")

fit_rw <- with_rw(imp, glm(D ~ A, subset = A > 2, family = binomial()))

pooled <- pool_rw(fit_rw)
pooled

```
