---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# `R/rw`: Robins-Wang variance estimation for multiple imputation

<!-- badges: start -->
[![R-CMD-check](https://github.com/LucyMcGowan/rw/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/LucyMcGowan/rw/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/LucyMcGowan/rw/graph/badge.svg)](https://app.codecov.io/gh/LucyMcGowan/rw)
<!-- badges: end -->

The purpose of the `rw` package is to compute [Robins-Wang variance estimates](https://doi.org/10.1093/biomet/87.1.113) for multiply imputed data analyses. The package implements RW-MICE for parametric chained-equation imputation using `method = "norm"` or `method = "logreg"`, and RW-PMM for predictive mean matching using `method = "pmmrw"`.

## Installation

This package requires the development version of mice with the `tasks` argument:

```r
remotes::install_github("amices/mice@dev")
```

Then you can install the development version of `rw` like so:

``` r
remotes::install_github("LucyMcGowan/rw")
```


## RW-MICE with parametric imputation

The first step is to impute your data using `mice`. When using the `mice` function, you must set the parameter `tasks = "train"` in order to pass the necessary parts to our subsequent functions. In this example, we use the `nhanes` data from the mice package and impute the incomplete variables `bmi`, `hyp`, and `chl` using normal working models. The variable `age` has no missing values.

```{r}
#| message: false
library(rw)
library(mice)
set.seed(1)
nhanes_scaled <- as.data.frame(scale(nhanes))

meth_norm <- make.method(nhanes_scaled)
meth_norm[] <- ""
meth_norm[c("bmi", "hyp", "chl")] <- "norm"

imp_norm <- mice(nhanes_scaled, 
                 method = meth_norm,
                 m = 5, 
                 tasks = "train",
                 print = FALSE)
```

Now, suppose we want to fit a model predicting `bmi` from `age` and `hyp`. The `with_rw` function allows this, where the first argument is the imputation object from the `mice` function above and the second is the model expression. Note that currently outcome models must be Gaussian or Binomial.

```{r}
fit_rw_norm <- with_rw(imp_norm, lm(bmi ~ age + hyp))
```

Finally, we can pool the results from the fit object and calculate the Robins-Wang variance using the `pool_rw` function

```{r}
pool_rw(fit_rw_norm)
```

Let's compare this result to using Rubin's rules.

```{r}
fit_rr_norm <- with(imp_norm, lm(bmi ~ age + hyp))
pool(fit_rr_norm) |>
  summary()
```


## RW-PMM with predictive mean matching

RW-PMM is the Robins-Wang variance estimator for analyses after predictive mean
matching. The `pmmrw` method uses the ordinary MICE hard donor draw and stores
the realized donor row. It additionally evaluates that donor under a smooth
all-observed-donor working law to construct the score required by `pool_rw()`.
The working law affects variance estimation only; it does not generate the
completed values.

Here we impute `bmi` using `age`, then analyze `bmi ~ age`.

```{r}
nhanes_pmm <- nhanes_scaled[c("age", "bmi")]

meth_pmm <- make.method(nhanes_pmm)
meth_pmm[] <- ""
meth_pmm["bmi"] <- "pmmrw"

set.seed(1)
imp_pmm <- mice(nhanes_pmm,
                method = meth_pmm,
                m = 5,
                tasks = "train",
                print = FALSE)
```

We can verify that `pmmrw` does not change the imputed values.

```{r}
meth_standard_pmm <- meth_pmm
meth_standard_pmm["bmi"] <- "pmm"

set.seed(1)
imp_standard_pmm <- mice(nhanes_pmm,
                         method = meth_standard_pmm,
                         m = 5,
                         print = FALSE)

identical(imp_standard_pmm$imp$bmi, imp_pmm$imp$bmi)
```

```{r}
fit_rw_pmm <- with_rw(imp_pmm, lm(bmi ~ age))
pool_rw(fit_rw_pmm)
```

Compare this with Rubin's rules for the same PMM imputations.

```{r}
fit_rr_pmm <- with(imp_pmm, lm(bmi ~ age))
pool(fit_rr_pmm) |>
  summary()
```

## Giganti & Shepherd Example


Below is an example to replicate the [Giganti & Shepherd (2020)](https://doi.org/10.1093/aje/kwaa153) results. This example uses the parametric `norm` and `logreg` imputation models and demonstrates the original Robins-Wang workflow; it is not a PMM correlation-adjusted example.

```{r example}
set.seed(1)
meth <- make.method(giganti_data)
meth[] <- ""
meth[c("A", "D")] <- c("norm", "logreg")

pred <- make.predictorMatrix(giganti_data)
pred[,] <- 0
pred["A", c("X1", "X2", "A.star", "D.star")] <- 1
pred["D", c("X1", "X2", "A.star", "D.star", "A")] <- 1

imp <- mice(giganti_data, m = 10, method = meth, predictorMatrix = pred,
            print = FALSE, tasks = "train")

fit_rw <- with_rw(imp, glm(D ~ A, subset = A > 2, family = binomial()))

pooled <- pool_rw(fit_rw)
pooled

```
